---
title: "PGA Golf Exploratory and Correlaton Analysis"
author: "Patrick McCarthy"
date: "7/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Libraries
```{r warning=FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(MASS)
```

Initial Data load
```{r}
golf <- read_csv("~/Documents/DS Capstone Project/Datasets/golf_fields_avgs_trim.csv")
```

```{r}
summary(golf)
```

Filtering out any player who didn't have more than 40 rounds (less than 1st Quartile) on the year.
```{r}
golf2 <- golf %>%
  filter(golf$rounds_played >= 40)
```

```{r}
summary(golf2)
```

Drop NA rows to get dataset with complete metrics for initial analysis
```{r}
golf3 <- golf2 %>%
  drop_na()
```

```{r}
summary(golf3)
```

Plot to show how driving distance has increased dramatically over the years and why previous modeling on metrics should be updated now that driving has changed
```{r}
ggplot(golf2 %>% group_by(year) %>% summarise(driving=mean(`Driving Distance_AVG`)), aes(x=year, y=driving)) +
  xlab("Year") +
  ylab("Average Driving Distance (yds)") +
  labs(title = "PGA Driving Distance over Time") +
  geom_point() +
  stat_smooth()
```
```{r}
ggplot(golf2 %>% group_by(year) %>% summarise(scoring=mean(`Scoring Average_Actual`)), aes(x=year, y=scoring)) +
  geom_point() +
  xlab("Year") +
  ylab("Scoring Average (Actual)") +
  labs(title = "Scoring Average Actual over Time") +
  stat_smooth()
```
```{r}
ggplot(golf2 %>% group_by(year) %>% summarise(gir=mean(`GIR_%`)), aes(x=year, y=gir)) +
  geom_point() +
  xlab("Year") +
  ylab("GIR %") +
  labs(title = "Greens in Regulation over Time") +
  stat_smooth()
```
```{r}
ggplot(golf2 %>% group_by(year) %>% summarise(driving_acc=mean(`Driving Accuracy_%`)), aes(x=year, y=driving_acc)) +
  geom_point() +
  xlab("Year") +
  ylab("Driving Accuracy %") +
  labs(title = "Driving Accuracy Percentage over Time") +
  stat_smooth()
```

CORRELATION TESTING
```{r}
library(corrplot)
#correlation dataset
golf_cor <- dplyr::select(golf3, -c(player,year, rounds_played, `Top 10 Finishes`, `First Place Finishes`, `Second Place Finishes`, `Third Place Finishes`))
res <- cor(golf_cor)

#correlogram
corrplot(res, type = "upper", tl.pos = "td", tl.cex = 0.5, order = "hclust", tl.col = "black", sig.level = .05, title = "Correlation Plot of PGA Tour Statistics")
```
Density Plot to Show Normality
```{r}
golf_cor %>%
  #keep(is.numeric) %>%
  gather() %>%
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()
```

```{r}
# ++++++++++++++++++++++++++++
# flattenCorrMatrix
# ++++++++++++++++++++++++++++
# cormat : matrix of the correlation coefficients
# pmat : matrix of the correlation p-values
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
```

Flattened Correlation Dataframe with P value
```{r}
library(Hmisc)
res2 <- rcorr(as.matrix(golf_cor))
flat_res <- flattenCorrMatrix(res2$r, res2$P)
```
Filtered Corr Table from above
```{r}
flat_res_trim <- flat_res %>%
  filter(flat_res$cor < -0.55 | flat_res$cor > 0.55)
```
Export Filtered Correlation Table
```{r}
write.csv(flat_res_trim, "~/Documents/DS Capstone Project/Datasets/flat_res_trim2.csv", row.names = FALSE)
```

Additional Testing
Correlation Test
```{r}
cor.test(golf_cor$`Scoring Average`, golf_cor$`Top 10 Finishes`, method = "pearson")
```

Scatter Plot between a Positively Correlated Pair of Variables
```{r}
library(ggpubr)
ggscatter(golf_cor, x = "Scoring Average", y = "Top 10 Finishes",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Scoring Average", ylab = "Top 10 Finishes")
```
```{r}
ggscatter(golf_cor, x = "GIR_%", y = "Scoring Average_Actual",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "GIR", ylab = "Scoring Average Actual")
```
```{r}
ggscatter(golf_cor, x = "Final Round Scoring Average_TOTAL ROUNDS", y = "Scrambling_PAR OR BETTER",
          add = "reg.line", conf.int = TRUE,
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Final Rounds", ylab = "Scrambling Par or Better")
```


